---
title: "Main"
author: "HADDOUCHE Théo, EL BAZ Avner"
date: "`r Sys.Date()`"
output: pdf_document
---
## Introduction
### Librairies

```{r}
library(tseries)
library(readr)
```

### Import et délimitation du dataset

```{r}

valeurs_mensuelles <- read_delim("valeurs_mensuelles.csv", delim = ";", col_types = cols())
vm <- valeurs_mensuelles[-c(1, 2, 3), ]
colnames(vm) <- c("date","value","code")  
vm <- vm[nrow(vm):1, ]
vm$value <- as.numeric(vm$value)
vm$diff <- c(NA,diff(vm$value))

str(vm)

head(vm)
```

## Part I :  The Data
1.  What does the chosen series represent ? (sector, potential data processing, logarithmic transformation, etc.)

La série représente la production

2. Transform the series to make it stationary if necessary (differentiate it, correct the deterministic
trend, etc.). Thoroughly justify your choices.

```{r}
serie_ts <- ts(vm$value, start = c(1990, 01), frequency = 12) 
diff_series <- ts(vm$diff, start = c(1990, 02), frequency = 12) 
# Test de Dickey-Fuller augmenté
diff_series <- na.omit(diff_series) 
adf_test <- adf.test(diff_series, alternative="stationary")
print(adf_test)
```

3. Graphically represent the chosen series before and after transforming it.

```{r}

plot(serie_ts, main="Série Temporelle", xlab="Temps", ylab="Valeur")
plot(diff_series, main="Série Temporelle Différenciée", xlab="Temps", ylab="Valeur")
```

2.  Transform the series to make it stationary if necessary (differentiate it, correct the deterministic trend, etc.). Thoroughly justify your choices.

```{r}



```
